---
title: "spdep LOSH calculations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## spdep LOSH function

```{r}
# From the documentation
library(spdep)
library(rgdal)

###########################################
# Prepare workspace

# Load in test data - Dnever housing data from https://geodacenter.github.io/data-and-lab//census_2010_excluded-variables/
# Downloaded as .zip, converted to .gpkg that is now downloaded from Github url
denver <- st_read("https://github.com/jeffcsauer/GSOC2020/raw/master/validation/data/denver/denver.gpkg")
  # Save as geopackage
    #outname <- "C:/Users/jeffe/Dropbox/GSOC2020/validation/data/denver/denver.gpkg"
    #denver_sp <- as_Spatial(denver)
    #writeOGR(denver_sp, dsn = outname, layer = "denver", driver = "GPKG")

############################################
# Create spatial weights

# Create a row-standardized neighbors list
denver_nb <- poly2nb(denver)
denver_lw <- nb2listw(denver_nb)
denver_lw$weights[1]

# Calculate losh measures
  #resLOSH <- as.data.frame(LOSH(denver$HU_RENTED, denver_lw))
  # With chisquare pvalues
  resLOSH <- as.data.frame(LOSH.cs(denver$HU_RENTED, denver_lw))
  resLOSH
    
  # Save to csv
    #write.csv(resLOSH, "C:/Users/jeffe/Dropbox/GSOC2020/validation/data/spdep_denver_losh.csv")

# Create a NON row-standardized neighbors list
denver_lw_ns <- nb2listw(denver_nb, style = "B")
denver_lw_ns$weights[1]

  # Calculate losh measures for non standardized list
  resLOSH_ns <- as.data.frame(LOSH.cs(denver$HU_RENTED, denver_lw_ns))
  
############################################
# Seems to be same result, why? 

  # Examine LOSH function
  LOSH
  
  # Note that authors first define Wi as rowsum of each list weight
  Wi <- vapply(denver_lw$weights, sum, FUN.VALUE = 0); Wi
  Wi_ns <- vapply(denver_lw_ns$weights, sum, FUN.VALUE = 0); Wi_ns
  
  # When calculating the mean, x_bar_i, they divide the spatial lag by the row sum of the spatial weights.
  # the x_bar_i values are then equal even when using non standardized row weights
  x_bar_i <- lag.listw(denver_lw, denver$HU_RENTED)/Wi; x_bar_i
  x_bar_i_ns <- lag.listw(denver_lw_ns, denver$HU_RENTED)/Wi_ns; x_bar_i_ns
  
    round(x_bar_i,3) == round(x_bar_i_ns,3)

  res_ei <- abs(denver$HU_RENTED - x_bar_i)^2; res_ei
  res_ei_ns <- abs(denver$HU_RENTED - x_bar_i_ns)^2; res_ei_ns
  
  # The only difference between the two is the denom_hi, but it doesn't seem to matter...
  denom_hi <- mean(res_ei) * Wi; denom_hi
  denom_hi_ns <- mean(res_ei_ns) * Wi_ns; denom_hi_ns
  
  # As the results are equal to each other!
  lag.listw(denver_lw_ns, res_ei_ns)
  
  res_Hi <- lag.listw(denver_lw, res_ei)/denom_hi; res_Hi
  res_Hi_ns <- lag.listw(denver_lw_ns, res_ei_ns)/denom_hi_ns; res_Hi_ns

  # Validate
  temp <- as.data.frame(LOSH.cs(denver$HU_RENTED, denver_lw))
  temp2 <- as.data.frame(LOSH.cs(denver$HU_RENTED, denver_lw_ns))

  